State of the Practice for Lattice Boltzmann Method Software

### 2022-04-24 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits to introduction

\begin{keyword}
	Lattice Boltzmann Method (LBM), research software, software engineering,
	software quality, Analytic Hierarchy Process
\end{keyword}

@@

A small sample of safety and mission-critical applications of LBM includes the
following: designing fuel cells \citep{ZhangEtAl2018}, modelling groundwater
flow \citep{AnwarAndSukop2009}, and analyzing the flow of blood in the
cardiovascular system \citep{SadeghiEtAl2022b, SadeghiEtAl2022,
SadeghiEtAl2020}.  As these examples illustrate, engineers and scientists use
LBM for tackling problems that impact such areas as environmental policy,
manufacturing, and health and safety. Given the critical applications of LBM,
users worry about with software qualities like reliability,
robustness, reproducibility, performance, correctness, and verifiability.  Since
their time is valuable, developers of LBM libraries have additional quality
concerns, including maintainability, reusability, and understandability.  With
these quality considerations in mind, and the potentially overwhelming number of choices for existing LBM libraries, the time is right to assess the state of the practice. Therefore, we analyze the current state of LBM
software development and compare it to the development practices employed for
research software in general.  We want to highlight success stories that LBM
developers can share amongst their community, and with the broader research
software community, while at the same time looking for areas for potential
future improvement.

We use ten research questions to structure and focus our discussion. Our
methodology \citep{SmithEtAl2021, SmithAndMichalski2022} is driven by the
questions. For each question below, we point to the section that contains our
answer.  We start with identifying the examples of LBM software where we have
access to the source code:

@@

We next wish to assess the representative software to determine how well they
apply current software development best practices.  By \emph{best practices}, we
mean methods, techniques, and tools that are generally believed to improve
software development, like testing and documentation.  We will categorize our
best practices around software qualities, as we will discuss in
Section~\ref{empiricalmeasures}.  At this point in the process, to remove
potential user/developer bias, we will base our assessment only on publicly
available artifacts, where artifacts are the documents, scripts, and code that we
find in a project's public repository. Example artifacts include requirements,
specifications, user manuals, unit test cases, system tests, usability tests,
build scripts, API (Application Programming Interface) documentation, READMEs,
license documents, process documents, and code. Following best practices does
not guarantee popularity, so we will also compare our ranking to how the user
community ranks the identified projects.

@@

\begin{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on LBM software projects?
	(Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from LBM compare to the pain points
	for research software in general? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For LBM
	developers what specific best practices address the pain points
	and software quality concerns? (Section~\ref{Sec_AddressConcerns})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practices could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}.
	(Section~\ref{Sec_Recommendations})
\end{enumerate}

@@

Compared with our previous methodology (as used in \citet{SmithEtAl2016} for
instance), measuring the application of best practices now also includes
repository based metrics, such as the number of files, number of lines of code,
percentage of issues that are closed, etc.  We rank the software with the
Analytic Hierarchy Process (AHP), using the quantitative data from the grading
template. After this, as another addition to our previous methodology, we
interview some development teams to understand the status of their
development processes. We next answer the research questions using the collected
quantitative and qualitative data (Sections~\ref{methodology}
to~\ref{Sec_AddressConcerns}).  Finally, we summarize the threats to validity
(Section~\ref{threats}) and the final conclusions and recommendations
(Section~\ref{conclusion}).

>>>>

\begin{keyword}
	Lattice Boltzmann Method (LBM), research software, software engineering,
	software quality, Analytic Hierarchy Process (AHP)
\end{keyword}

@@

A small sample of safety and mission-critical applications of LBM includes the
following: designing fuel cells \citep{ZhangEtAl2018}, modelling groundwater
flow \citep{AnwarAndSukop2009}, and analyzing the flow of blood in the
cardiovascular system \citep{SadeghiEtAl2022b, SadeghiEtAl2022,
SadeghiEtAl2020}.  As these examples illustrate, engineers and scientists use
LBM for tackling problems that impact such areas as environmental policy,
manufacturing, and health and safety. Given the critical applications of LBM,
users worry about software qualities like reliability, robustness,
reproducibility, performance, correctness, and verifiability.  Since their time
is valuable, developers of LBM libraries have additional quality concerns,
including maintainability, reusability, and understandability.  With these
quality considerations in mind, and the potentially overwhelming number of
choices for existing LBM libraries, the time is right to assess the state of the
practice. Therefore, we analyze the current state of LBM software development
and compare it to the development practices employed for research software in
general.  We want to highlight success stories that LBM developers can share
amongst their community, and with the broader research software community, while
at the same time looking for areas for potential future improvement.

We use ten research questions to structure and focus our discussion. Our
methodology \citep{SmithEtAl2021, SmithAndMichalski2022} is driven by these
questions. For each question below, we point to the section that contains our
answer.  We start with identifying the examples of LBM software where we have
access to the source code:

@@

We next wish to assess the representative software to determine how well they
apply current software development best practices.  By \emph{best practices}, we
mean methods, techniques, and tools that are generally believed to improve
software development, like testing and documentation.  We categorize our best
practices around software qualities (Section~\ref{softwarequalities}).  At this
point in the process, to remove potential user/developer bias, we base our
assessment only on publicly available artifacts, where artifacts are the
documents, scripts, and code that we find in a project's public repository.
Example artifacts include requirements, specifications, user manuals, unit test
cases, system tests, usability tests, build scripts, API (Application
Programming Interface) documentation, READMEs, license documents, process
documents, and code. Following best practices does not guarantee popularity, so
we also compare our ranking to how the user community ranks the identified
projects.

@@

\begin{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on LBM software projects?
	(Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from LBM compare to the pain points
	for research software in general? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For LBM
	developers what specific best practices address their pain points
	and software quality concerns? (Section~\ref{Sec_AddressConcerns})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practices could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}.
	(Section~\ref{Sec_Recommendations})
\end{enumerate}

@@

Compared with our previous methodology (as used in \citet{SmithEtAl2016} for
instance), measuring the application of best practices now also includes
repository based metrics, such as the number of files, number of lines of code,
percentage of issues that are closed, etc.  We rank the software with the
Analytic Hierarchy Process (AHP), using the quantitative data from the grading
template. After this, as another addition to our previous methodology, we
interview some development teams to understand the status of their
development processes. We next answer the research questions using the collected
quantitative and qualitative data (Sections~\ref{methodology}
to~\ref{Sec_Recommendations}).  Finally, we summarize the threats to validity
(Section~\ref{threats}) and the final conclusions and recommendations
(Section~\ref{conclusion}).
